---
title: "TNBC Pharmacogenomic analysis"
---
This section initiallizes the R environment and loads custom R functions throughput the pipeline.
```{r message=FALSE, warning=FALSE, include=FALSE}
rm(list=ls())
library("glmnet")
library("doMC")
library("reticulate")
library("plotly")
library("matrixStats")
library("ComplexHeatmap")
library("seriation")
library("dplyr")
library("glmnet")
library("Hmisc")

plot_AvP = function(AVP_mat){
  library("ggpubr")
  library("ModelMetrics")
  
  temp = as.data.frame(AVP_mat[,c('actual','predicted')])
  temp[] <- lapply(temp, function(x) {if(is.factor(x)) as.numeric(as.character(x)) else x})
  sapply(temp, class)
  
  sp = ggscatter(temp,
                  x="actual",
                  y="predicted",
                  color = "lightgray",
                  add = "reg.line",
                  add.params = list(color = "black"),
                  conf.int = TRUE,
                  cor.coef = TRUE,
                  cor.method = "pearson",)
  
  sp + geom_density_2d()
}
```

This section initiallizes the Python environment and loads custom Python functions throughput the pipeline.
Portions of this code are taken from: https://github.com/KnowEnG/ProGENI (Paper: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5554409/)
```{python}
import os
import sys
import argparse
import warnings
import numpy as np
from numpy import mean
import pandas as pd
import scipy.stats as ss
from scipy import sparse
from scipy.stats.mstats import zscore
from scipy.stats.mstats import gmean
from scipy.sparse import SparseEfficiencyWarning
from sklearn.preprocessing import normalize

def is_number(num):
  try:
      float(num)
      return True
  except ValueError:
      return False

###############################################################################
def spread_match_network(expr_df_in, node_names_in):
  node_names_in_set = set(node_names_in)
  gene_names_in_set = set(expr_df_in.columns.values)
  
  nodes_genes_intersect = sorted(list(gene_names_in_set & node_names_in_set))
  nodes_minus_genes = sorted(list(node_names_in_set - gene_names_in_set))
  genes_minus_nodes = sorted(list(gene_names_in_set - node_names_in_set))
  
  genes_names_out = nodes_genes_intersect + genes_minus_nodes
  nodes_names_out = nodes_genes_intersect + nodes_minus_genes
  expr_df_out = expr_df_in[genes_names_out]
  return(expr_df_out, nodes_names_out, nodes_genes_intersect)

###############################################################################
def rwr_matrix(node_names, network_matrix, restart_matrix, restart_prob, max_iter, tolerance):
  no_restart_prob = 1 - restart_prob
  init_prob = 1/len(node_names)
  # Create the vector of probabilities for the nodes
  steady_prob_old = np.empty(np.shape(restart_matrix))
  steady_prob_old.fill(init_prob)
  residual = 100
  num_iter_tmp = 0
  while (residual > tolerance) and (num_iter_tmp < max_iter):
      steady_prob_new = (sparse.csr_matrix.dot(steady_prob_old, network_matrix) \
                          * no_restart_prob + restart_prob \
                          * restart_matrix)
      residual = max(abs(steady_prob_new - steady_prob_old).sum(axis=1))
      print('iteration = ', num_iter_tmp)
      num_iter_tmp += 1
      steady_prob_old = steady_prob_new.copy()
  return(num_iter_tmp, residual, steady_prob_new)

def import_network(address_net, delimiter):
  """
  Imports the network and generates a dataframe.
  Input:
      address_net: The address of the network
      delimiter: The delimiter used to import the network
  """
  default_column_headers = ['n_alias_1', 'n_alias_2', 'weight', 'type']
  # Step 1: Read the input
  with open(address_net, 'r') as fin:
      # Check whether the first line is data or headers
      first_line = fin.readline().strip()
      fin.seek(0)   #go back to the beginning of the file
      fields = first_line.split(sep=delimiter)
      # data
      if is_number(fields[2]):
          net_df = pd.read_csv(fin, sep=delimiter, names=default_column_headers)
      # headers
      else:
          net_df = pd.read_csv(fin, sep=delimiter, header=0)
  
  # Get the column headers
  node1 = net_df.columns[0]
  node2 = net_df.columns[1]
  weight = net_df.columns[2]
  #t = net_df.columns[3]
  
  # Get the unique nodes -- the first two columns of the input data,
  nodes1 = net_df.iloc[:, 0]
  nodes2 = net_df.iloc[:, 1]
  nodes = set(nodes1) | set(nodes2)
  node_names = sorted(nodes)
  num_nodes = len(node_names)
  
  # Output some info about the input data
  print("Number of rows in the network file:", len(net_df))
  print("Number of unique nodes in the network:", num_nodes)
  
  return(node_names, net_df, node1, node2, weight)

###############################################################################
def gen_network_matrix(num_nodes, net_df, node1, node2, weight, node2index):
  """Generates network adjacency matrix and normalizes it"""
  # Transform the first two columns of the DataFrame -- the nodes -- to their indexes
  net_df[node1] = net_df[node1].apply(lambda x: node2index[x])
  net_df[node2] = net_df[node2].apply(lambda x: node2index[x])
  # Create the sparse matrix
  network_matrix = sparse.csr_matrix((net_df[weight].values,(net_df[node1].values, net_df[node2].values)),shape=(num_nodes, num_nodes), dtype=float)
  # Make the ajdacency matrix symmetric
  network_matrix = (network_matrix + network_matrix.T)
  network_matrix.setdiag(0)
  # Normalize the rows of network_matrix because we are multiplying vector by matrix (from left)
  network_matrix = normalize(network_matrix, norm='l1', axis=1)
  return(net_df, network_matrix)
```

Set working directory and read in the experimental data frames
```{r message=FALSE, warning=FALSE}
library("readxl")
Base_dir = "Q:\\MDA_Helen PiwnicaWorms\\TNBC_PGx"
knitr::opts_knit$set(root.dir = Base_dir)
setwd(Base_dir)

Drug_all <- read.csv("./data/Drug_AUC.csv",row.names=1)
Drug_meta <- read.csv("./data/Drug_metadata.csv",row.names=1)
RNA_all <- read.csv("./data/RNAseq_COMBAT.csv",row.names=1)
Meta_all <- read_excel("./data/PDX_metadata.xlsx")
PPI = "./data/STRING.csv"

row.names(Meta_all) = Meta_all$PDX_ID
```

```{r}
library("matrixStats")
RNA_means = rowMeans(as.matrix(RNA_all))
RNA_stdDev = rowSds(as.matrix(RNA_all))
plot(RNA_means,RNA_stdDev)

RNA_filter = which(RNA_means >= .5)
RNA_filtered = RNA_all[RNA_filter,]
```

```{r, warning=FALSE}
library("RColorBrewer")

nsamples <- ncol(RNA_filtered)
col <- brewer.pal(nsamples, "Paired")
par(mfrow=c(1,2))
plot(density(as.matrix(RNA_filtered[,1])), col=col[1], lwd=2, ylim=c(0,1), las=2, main="", xlab="")
title(main="Filtered data", xlab="RNAseq")
#abline(v=lcpm.cutoff, lty=3)
for (i in 2:nsamples){
  den <- density(as.matrix(RNA_filtered[,i]))
  lines(den$x, den$y, col=col[i], lwd=2)
}

boxplot(RNA_filtered[,], las=2, col=col, main="")
title(main="Filtered data",ylab="RNAseq")
```

This section used RWR network-diffusion algorithm to smooth gene expression data using the human STRING-PPI network.
```{python}
#n_rcg = 100
restart_prob_trans = 0.5
restart_prob = 0.5
tolerance = 1e-8
max_iter = 100
n_boot = 1
percent_boot = 100/100

address_net = r["PPI"]

expr_df = r["RNA_all"].T
delimiter_net = ','
(node_names, net_df, node1, node2, weight) = import_network(address_net, delimiter_net)

###############################################################################
# Reorder gene column names of expression spreadsheet and gene node names of network
# such that both name lists start with the intersection of the two list ordered alphabetically
(expr_df, node_names, nodes_genes_intersect) = spread_match_network(expr_df, node_names)
ccl_names = expr_df.axes[0].tolist()


expr_all = zscore(expr_df.values, axis=0)
gene_names = list(expr_df.columns.values)
if len(set(gene_names)) != len(gene_names):
    sys.exit("Duplicate gene names!")
    
node2index = {node_names[i]:i for i in range(len(node_names))}
index2node = node_names.copy()

num_nodes = len(node_names)
print('Number of genes in the expression file:', len(gene_names))

###############################################################################
#Perform network transformation on the gene expression matrix
(net_df, network_matrix) = gen_network_matrix(num_nodes, net_df, node1, node2, weight, node2index)
restart_matrix = np.eye(len(node_names), len(nodes_genes_intersect)).T
print('Obtaining the network transformed gene expressions:')

(num_iter, residual, gene_similarity_smooth) = rwr_matrix(node_names, network_matrix, restart_matrix, restart_prob_trans, max_iter, tolerance)
gene_similarity_smooth = gene_similarity_smooth[:, 0:np.size(gene_similarity_smooth, axis=0)]
gene_similarity_smooth = normalize(gene_similarity_smooth, norm='l1', axis=1)
expr_all = zscore(expr_all[:, 0:len(nodes_genes_intersect)].dot(gene_similarity_smooth.T), axis=0)
```

```{r include=FALSE, paged.print=TRUE}
#Read smooetheed RWR df back in from python env
RNA_filtered_RWR = t(py$expr_all)
colnames(RNA_filtered_RWR) = py$ccl_names
row.names(RNA_filtered_RWR) = py$nodes_genes_intersect
```

We aimed to identify specific drug mechanisms and functional targets that are associated with chemo-resistant TNBC. Accordingly, we filter out longitudinal matched sets from "orphan" treatment niave PDX models.
```{r warning=FALSE}
pairsOnly = T
cellLine_drug = colnames(Drug_all)
drugnames = row.names(Drug_all)

cellLine_RNA = colnames(RNA_filtered_RWR)
genenames = row.names(RNA_filtered_RWR)

all_samples = intersect(cellLine_drug,cellLine_RNA)

Pairs = Meta_all %>% count(PiD)
Pairs = Pairs[which(Pairs$n >= 2),]
Meta_pairsOnly = Meta_all[which(Meta_all$PiD %in% Pairs$PiD),]
row.names(Meta_pairsOnly) = Meta_pairsOnly$PDX_ID

if (pairsOnly){
  train_samples = intersect(all_samples,Meta_pairsOnly$PDX_ID)
} else{train_samples = all_samples}
train_samples

tmp = setdiff(colnames(RNA_filtered_RWR),train_samples)
test_samples = intersect(colnames(Drug_all),tmp)
unknown_samples = setdiff(tmp,colnames(Drug_all))

with_held_RNA = RNA_filtered_RWR[,test_samples]
with_held_Drug = Drug_all[,test_samples]
with_held_common = intersect(colnames(with_held_RNA),colnames(with_held_Drug))

with_held_RNA = with_held_RNA[,with_held_common]
with_held_Drug = with_held_Drug[,with_held_common]
rm("tmp")
```

```{r message=FALSE, warning=FALSE}
library("TNBC.CMS")
TNBCtype = predictCMS(RNA_all)
table(TNBCtype)
```

```{r}
resultGES <- computeGES(expr = RNA_all, pred = TNBCtype,rnaseq = TRUE)
```

```{r}
resultGSVA <- performGSVA(expr = RNA_all, pred = TNBCtype,gene.set = NULL)
```

```{r}
resultDS <- computeDS(expr = RNA_all, pred = TNBCtype)
```

We then qualitatively visualized the conversion of TNBC subtypes over time in those matched pairs. 
```{r echo=FALSE}
library(tidyr)
library(dplyr)
library(networkD3)

tmp = as.data.frame(Meta_pairsOnly)[,c("PiD","TNBCtype","Time")]
tmp = reshape(tmp, direction = "wide", idvar = "PiD", timevar = "Time")
colnames(tmp) = c('ID','Pre','Mid','Pst')
tmp = tmp[,c('Pre','Mid','Pst')]

links <-
  tmp %>% 
  mutate(row = row_number()) %>%  # add a row id
  pivot_longer(-row, names_to = "col", values_to = "source") %>%  # gather all columns
  mutate(col = match(col, names(tmp))) %>%  # convert col names to col ids
  mutate(source = paste0(source, '_', col)) %>%  # add col id to node names
  group_by(row) %>%
  mutate(target = lead(source, order_by = col)) %>%  # get target from following node in row
  ungroup() %>% 
  filter(!is.na(target)) %>%  # remove links from last column in original data
  group_by(source, target) %>% 
  summarise(value = n(), .groups = "drop")  # aggregate and count similar links

# create nodes data frame from unque nodes found in links data frame
nodes <- data.frame(id = unique(c(links$source, links$target)),
                    stringsAsFactors = FALSE)

# remove column id from node names
nodes$name <- sub('_[0-9]*$', '', nodes$id)
nodes[nodes=="NA"] = NA

# create node ids in links data to the 0-based index of the nodes in the nodes data frame
links$source_id <- match(links$source, nodes$id) - 1
links$target_id <- match(links$target, nodes$id) - 1

sn = sankeyNetwork(Links = links,
                         Nodes = nodes,
                         Source = 'source_id',
                         Target = 'target_id',
                         Value = 'value',
                         NodeID = 'name',
                         fontSize = 12,
                         iterations = 5)
sn
```

Next, we evaluated the similarity of the matched pairs to determine whether individual patients or time points tended to associate with one another. From this analysis we found that individual patients tended to co-cluster over time points. Thus recapitulating a strong degree of patient heterogeneity.
```{r echo=FALSE}
library("dendextend")
library("ComplexHeatmap")

CCL_corr = cor(as.matrix(RNA_filtered_RWR[,train_samples]),method = c("pearson"))
df_anno = Meta_all[train_samples,]%>%select("Time","PiD")

K = 9
hc = hclust(dist(CCL_corr))
clus = cutree(hc,k=K)

df_anno_clust = cbind(df_anno,clus)
chisq.test(df_anno_clust$PiD, df_anno_clust$clus, correct=FALSE)
table(df_anno_clust$PiD, df_anno_clust$clus)

row_dend = as.dendrogram(hc)
row_dend = color_branches(row_dend, k = K)

hm_anno = HeatmapAnnotation(df = as.data.frame(df_anno),
                            annotation_name_side = "left",
                            col=list(
                              Time=c("PRE"="grey80",
                                     "MID"="grey40",
                                     "POST"="grey0"
                                          ),
                              PiD =c("2"="yellow3",
                                     "3"="wheat3",
                                     "11"="violetred3",
                                     "14"="turquoise3",
                                     "16"="tomato3",
                                     "18"="thistle3",
                                     "19"="tan3",
                                     "20"="steelblue3",
                                     "21"="springgreen3",
                                     "22"="snow3"
                                     )
                              )
                            )
HM = Heatmap(CCL_corr,
             name = "Pearson",
             cluster_rows = row_dend,
             cluster_columns = row_dend,
             top_annotation = hm_anno,
             show_row_names = FALSE)
HM
```

```{r echo=FALSE, warning=FALSE}
library(apcluster)
library(dplyr)
library(crosstable)

apclust = apcluster(s = CCL_corr,
                    p=NA,
                    q=NA,
                    maxits=1000,
                    convits=100,
                    lam=0.9,
                    includeSim=FALSE,
                    details=FALSE,
                    nonoise=FALSE,
                    seed=NA)
show(apclust)

clusters <- attr(apclust, "clusters")
centers <- attr(apclust, "exemplars")
nclust <- length(centers)

res <- NULL
for (i in 1:nclust) {
	mems <- clusters[[i]]
	clus <- rep(i, length(mems))
	ctr <- ifelse(mems == centers[i], i, 0)
	df <- data.frame(mems, clus, ctr)
	res <- rbind(res, df)
}

df_anno_clust = cbind(df_anno,res)
chisq.test(df_anno_clust$PiD, df_anno_clust$clus, correct=FALSE)
table(df_anno_clust$PiD, df_anno_clust$clus)
```

We then perform differential gene expression analysis that accounted for the individual patient matched pairs. 
```{r}
library("edgeR")
metadata = Meta_all[train_samples,]
metadata[] <- lapply(metadata, factor)

design = model.matrix(~0+Time+PiD,metadata)
DEG_fit = lmFit(RNA_all[,train_samples],design)
cm = makeContrasts(Mid_v_Pre=TimeMID-TimePRE,levels=design)
DEG_fit_cm = contrasts.fit(DEG_fit,cm)
DEG_fit_cm_eBayesfit = eBayes(DEG_fit_cm,trend = FALSE,robust = FALSE)
DEG_top = topTable(DEG_fit_cm_eBayesfit,adjust.method="fdr", number=Inf)

DEG_top
```


```{r message=FALSE, warning=FALSE}
library("plotly")
fig <- plot_ly(data = DEG_top, x = ~logFC, y = ~-1*log10(P.Value),color = ~logFC, colors="RdBu")%>%
  layout(xaxis = list(title = "log(FC)"),yaxis = list(title = "-log(p-value)")) %>%
  add_trace(text=row.names(DEG_top))
fig
```

```{r}
tmp = DEG_top[which(DEG_top$P.Value <= 0.05),]
up = as.integer(length(which(tmp$logFC > 0)))
down =  as.integer(length(which(tmp$logFC <= 0)))
print(paste("Total DEG: ", up+down))
print(paste("Sigificantly up-reglated genes: ", up))
print(paste("Sigificantly down-reglated genes: ", down))

rm("tmp","up","down")
```

```{r}
library("pathfindR")
pfR_input = DEG_top[,c("logFC","P.Value"),]
Gene_symbol = row.names(pfR_input)
pfR_input = cbind(Gene_symbol,pfR_input)

name = "20220916_DEG_pathfindR"

unlink(name, recursive = TRUE)

DEG_pathfindR <- run_pathfindR(pfR_input,
                               iterations = 10,
                               p_val_threshold = 0.05,
                               search_method = "GR",
                               plot_enrichment_chart = FALSE,
                               visualize_enriched_terms = FALSE,
                               output_dir = name,
                               silent_option = TRUE)
DEG_pathfindR
```

```{r fig.height=12, fig.width=12}
enrichment_chart(
  DEG_pathfindR,
  top_terms = Inf,
  plot_by_cluster = TRUE,
  num_bubbles = 5,
  even_breaks = TRUE
)
```
```{r}
library("visNetwork")

result_df = DEG_pathfindR
result_df <- result_df[order(result_df$lowest_p, decreasing = FALSE), ]

ID_column <- "Term_Description"

edges <- data.frame()
all_up = c()
all_down = c()
for (i in base::seq_len(nrow(result_df))) {
  up_genes <- unlist(strsplit(result_df$Up_regulated[i],", "))
  all_up = unique(c(all_up,up_genes))
  down_genes <- unlist(strsplit(result_df$Down_regulated[i],", "))
  all_down = unique(c(all_down,down_genes))
  genes <- c(up_genes, down_genes)
  for (gene in genes) {
    edges <- rbind(edges, data.frame(Term = result_df[i, ID_column], Gene = gene))
  }
}



pathways = as.data.frame(unique(edges$Term))
colnames(pathways) = "id"
pathways[,"color"] = "grey"
pathways[,"group"] = "Pathway"
pathways[,"shape"] = "square"
pathways[,"font.size"] = 30

all_up = as.data.frame(all_up)
colnames(all_up) = "id"
all_up[,"color"] = "green"
all_up[,"group"] = "Up_regulated"
all_up[,"shape"] = "circle"
all_up[,"font.size"] = 20

all_down = as.data.frame(all_down)
colnames(all_down) = "id"
all_down[,"color"] = "darkred"
all_down[,"group"] = "Down_regulated"
all_down[,"shape"] = "circle"
all_down[,"font.size"] = 20

nodes = rbind(pathways,all_up,all_down)


colnames(edges) = c("from","to")

rm("pathways","all_up","all_down")

graph = visNetwork(nodes, edges,width='100%',height=2000)%>%
  visNodes(color = list(inherit=FALSE))%>%
  visEdges(color = list(inherit=FALSE))%>%
  visOptions(highlightNearest = list(enabled = T, degree = 2, hover = T)) %>%
  visInteraction(multiselect = TRUE) %>%
  visLayout(randomSeed = 123)%>%
  visGroups(groupname = "Pathway", color = "grey") %>%
  visGroups(groupname = "Up_regulated", color = "green") %>%
  visGroups(groupname = "Down_regulated", color = "darkred") %>%
  visLegend(position = "right")
visSave(graph, file="DEG_term_enrichment_network.html", selfcontained = TRUE, background = "white")
graph
```

```{r fig.height=4, fig.width=14}
term_gene_heatmap(DEG_pathfindR,
                  use_description = TRUE,
                  num_terms = 25,
                  low = "red",
                  mid = "black",
                  high = "green")
```

```{r fig.height=12, fig.width=18}
term_gene_graph(DEG_pathfindR,
                use_description = TRUE,
                num_terms = Inf,
                node_size = "num_genes",
                layout = "stress")
```

Next, we looked at differential drug susceptibilities identified by the high throughput chemical screening assays. 
```{r}
library(matrixStats)
drug_range = rowRanges(as.matrix(Drug_all),useNames = TRUE)
drug_range[drug_range < 0] <- 0
drug_range = drug_range[,2]-drug_range[,1]

drug_filter = which(drug_range >= 0.5)
```

From the differential gene expression analysis comparing matched pairs, we observed enrichment of cell cycle related genes. This observation is consistent with the growth analysis performed on the ex vivo cultures from the high throughput screens, which show increases in growth rates when comparing the postNACT to the preNACT PDX models. It is also well established that alterations in growth rate and confound PGx associations. Accordingly, in this section we perfromed a correlative analysis of the growth data to the AUC activities.
```{r}
ord = colnames(Drug_all)
gr = Meta_all[ord,]
test = cor(gr$Log2_FC_3day,t(Drug_all))
r2 = test^2

#get list of correlated and variable drug names
growth_corr = names(r2[,which(r2 > 0.25)])
variable_drugs = names(drug_filter)
x = list(Growth_correlated_drugs = growth_corr,
         Variable_drug_response = variable_drugs)

library(ggvenn)
ggvenn(x,
       fill_color = c("#0073C2FF","#CD534CFF"),
       stroke_size = 0.5, set_name_size = 4)
```
This is the list of variable drugs that are potentially confounded by the rate of growth
```{r}
intersect(growth_corr,variable_drugs)
```
And this is the list of drugs that are not correlated with growth.
```{r}
drug_filter = variable_drugs[!(variable_drugs %in% growth_corr)]
drug_filter
```

```{r}
library("ComplexHeatmap")
#CCL_corr = cor(as.matrix(t(scale(t(Drug_all[drug_filter,train_samples]),center=TRUE, scale=TRUE))),method = c("pearson"))
CCL_corr = Drug_all
df_anno = Meta_all[colnames(CCL_corr),]%>%select("Time","TNBCtype")

library("dendextend")
row_dend = as.dendrogram(hclust(dist(Drug_all)))
row_dend = color_branches(row_dend, k = 4)

hm_anno = HeatmapAnnotation(df = as.data.frame(df_anno),
                            annotation_name_side = "left",
                            col=list(
                              Time=c("PRE"="darkolivegreen",
                                          "MID"="darkorange",
                                          "POST"="darkorchid"
                                           ),
                               PiD =c("2"="yellow3",
                                      "3"="wheat3",
                                      "11"="violetred3",
                                      "14"="turquoise3",
                                      "16"="tomato3",
                                      "18"="thistle3",
                                      "19"="tan3",
                                      "20"="steelblue3",
                                      "21"="springgreen3",
                                      "22"="snow3"
                                     )
                              )
                            )
HM = Heatmap(CCL_corr,
             name = "Pearson",
             cluster_rows = row_dend,
             top_annotation = hm_anno,
             show_row_names = FALSE)
HM
```

```{r}
library("matrixStats")
library("ggplot2")

MEAN_AUC = rowMeans(Drug_all)

hc = hclust(dist(Drug_all))
Cluster_ID = cutree(hc,k=4)

df = as.data.frame(cbind(Cluster_ID,MEAN_AUC))
df$Cluster_ID = as.factor(df$Cluster_ID)

ggplot(df, aes(x=Cluster_ID, y=MEAN_AUC, fill=Cluster_ID)) + geom_violin(trim=FALSE) + geom_boxplot(width=0.1)
```

```{r}
library("sunburstR")
library("d3r")

dat = Drug_meta[names(which(clus %in% c(3,4))), ] %>% select("Class","Target")
dat[,"Drug_name"] = row.names(dat) 
dat = dat[which(dat$Class != ""),]

dat = data.frame(lapply(dat, function(v) {
  return(toupper(v))
}))

df = dat %>% dplyr::count(Class,Target,Drug_name)
colnames(df)[ncol(df)] <- "size"

tree <- d3_nest(df, value_cols = "size")
sb1 <- sunburst(tree, width="100%", height=400,count = T, sortFunction = htmlwidgets::JS(
      "
    function(a,b) {
      // sort by count descending
      //   unlike the other example using data.name, value is at the top level of the object
      return b.value - a.value
    }
    "
)
  
)
htmltools::save_html(sb1, file = "./Clus4_panActive.html")
sb1
```


```{r}
library("dendextend")
library("ComplexHeatmap")

CCL_corr = cor(as.matrix(t(scale(t(Drug_all[drug_filter,train_samples]),center=TRUE, scale=TRUE))),method = c("pearson"))
df_anno = Meta_all[train_samples,]%>%select("Time","PiD")

K = 7
hc = hclust(dist(CCL_corr))
clus = cutree(hc,k=K)

df_anno_clust = cbind(df_anno,clus)
chisq.test(df_anno_clust$PiD, df_anno_clust$clus, correct=FALSE)
table(df_anno_clust$PiD, df_anno_clust$clus)

row_dend = as.dendrogram(hc)
row_dend = color_branches(row_dend, k = K)

hm_anno = HeatmapAnnotation(df = as.data.frame(df_anno),
                            annotation_name_side = "left",
                            col=list(
                              Time=c("PRE"="grey80",
                                          "MID"="grey40",
                                          "POST"="grey0"
                                          ),
                              PiD =c("2"="yellow3",
                                     "3"="wheat3",
                                     "11"="violetred3",
                                     "14"="turquoise3",
                                     "16"="tomato3",
                                     "18"="thistle3",
                                     "19"="tan3",
                                     "20"="steelblue3",
                                     "21"="springgreen3",
                                     "22"="snow3"
                                     )
                              )
                            )
HM = Heatmap(CCL_corr,
             name = "Pearson",
             top_annotation = hm_anno,
             show_row_names = FALSE)
HM
```

```{r}
library(apcluster)
library(dplyr)
library(crosstable)

apclust = apcluster(s = CCL_corr,
                    p=NA,
                    q=NA,
                    maxits=1000,
                    convits=100,
                    lam=0.9,
                    includeSim=FALSE,
                    details=FALSE,
                    nonoise=FALSE,
                    seed=NA)
show(apclust)

clusters <- attr(apclust, "clusters")
centers <- attr(apclust, "exemplars")
nclust <- length(centers)

res <- NULL
for (i in 1:nclust) {
	mems <- clusters[[i]]
	clus <- rep(i, length(mems))
	ctr <- ifelse(mems == centers[i], i, 0)
	df <- data.frame(mems, clus, ctr)
	res <- rbind(res, df)
}

df_anno_clust = cbind(df_anno,res)
chisq.test(df_anno_clust$PiD, df_anno_clust$clus, correct=FALSE)
table(df_anno_clust$PiD, df_anno_clust$clus)
```

```{r}
metadata = Meta_all[train_samples,]
metadata[] <- lapply(metadata, factor)

design = model.matrix(~0+Time+PiD,metadata)
Rx_fit = lmFit(Drug_all[drug_filter,train_samples],design)
Rx_fit_eBayesfit = eBayes(Rx_fit,trend = TRUE,robust = TRUE)
Rx_top = topTable(Rx_fit_eBayesfit,coef = "TimePRE" ,adjust.method="fdr",p.value = 0.01, number=Inf)
#Rx_top = topTable(Rx_fit_eBayesfit,coef = "TimePRE" ,adjust.method="fdr", number=Inf)
#Rx_top = topTable(Rx_fit_eBayesfit,coef = "TimeX0" ,adjust.method="fdr", number=Inf)
# cm = makeContrasts(Mid_v_Pre=TimeX1-TimeX0,levels=design)
# Rx_fit_cm = contrasts.fit(Rx_fit,cm)
# Rx_fit_cm_eBayesfit = eBayes(Rx_fit_cm,trend = TRUE,robust = TRUE)
# Rx_top = topTable(Rx_fit_cm_eBayesfit,adjust.method="fdr", number=Inf)
Rx_top
```

```{r fig.height=30, fig.width=42}
Mid_v_Pre = c()
Pst_v_Pre = c()
for (i in 1:nrow(Drug_all)){
  tmp1 = t(Drug_all[i,train_samples])
  colnames(tmp1) = "Active"
  
  tmp2 = Meta_all[train_samples,]
  tmp3 = cbind(tmp1,tmp2)
  tmp4 = tmp3[,c("Active","Time","PiD")]
  
  tmp5 = reshape(tmp4, direction = "wide", idvar = "PiD", timevar = "Time")
  tmp6 = tmp5$Active.MID-tmp5$Active.PRE
  tmp7 = tmp5$Active.POST-tmp5$Active.PRE
  names(tmp6) = tmp5$PiD
  names(tmp7) = tmp5$PiD
  
  Mid_v_Pre = rbind(Mid_v_Pre,tmp6)
  Pst_v_Pre = rbind(Pst_v_Pre,tmp7)
}

rm(list = c("tmp1","tmp2","tmp3","tmp4","tmp5","tmp6","tmp7"))

rownames(Mid_v_Pre) = row.names(Drug_all)
Mid_v_Pre = Mid_v_Pre[ , colSums(is.na(Mid_v_Pre)) == 0]
Mid_v_Pre = Mid_v_Pre[rownames(Rx_top),]

rownames(Pst_v_Pre) = row.names(Drug_all)
Pst_v_Pre = Pst_v_Pre[ , colSums(is.na(Pst_v_Pre)) == 0]
Pst_v_Pre = Pst_v_Pre[rownames(Rx_top),]

group_df = as.data.frame(Drug_meta[row.names(Rx_top),] %>% select("Class"),use_raster = TRUE, raster_by_magick = TRUE)
anno_df = as.data.frame(Drug_meta[row.names(Rx_top),] %>% select("Target"),use_raster = TRUE, raster_by_magick = TRUE)
##dAUC = rowMeans(cbind(Mid_v_Pre,Pst_v_Pre))
anno_df = cbind(X1_v_X0=rowMeans(Mid_v_Pre),XS_v_X0=rowMeans(Pst_v_Pre),anno_df)
anno_txt = anno_text(anno_df$Target, which = "row")

library(circlize)
col_fun = colorRamp2(c(-0.33,0,0.33), c("blue","white","red"))
#col_fun(seq(-10, 10))

rAnno = rowAnnotation(df = anno_df,
                      col=list(X1_v_X0=col_fun,XS_v_X0=col_fun), 
                      width = ncol(anno_df)*unit(10, "mm")
                      )


hm1 = Heatmap(Mid_v_Pre,
              width = ncol(Mid_v_Pre)*unit(20, "mm"),
              name = "dAUC",
              clustering_method_rows = "ward.D2",
              column_title = "MID-PRE",
              column_title_gp=gpar(fontsize=20,fontface = "bold"),
              column_names_gp = gpar(fontsize = 24),
              row_split = group_df$Class,
              gap = unit(2,"mm"),
              row_title_rot = 0,
              row_title_gp=gpar(fontsize=24,fontface = "bold"),
              heatmap_legend_param = list(direction = "vertical"),
              col = col_fun,
              use_raster = TRUE, raster_by_magick = TRUE)

hm2 = Heatmap(Pst_v_Pre,
              width = ncol(Pst_v_Pre)*unit(20, "mm"),
              name = "dAUC",
              column_title = "POST-PRE",
              column_title_gp=gpar(fontsize=20,fontface = "bold"),
              column_names_gp = gpar(fontsize = 24),
              heatmap_legend_param = list(direction = "vertical"),
              row_labels = anno_df$Target,
              col = col_fun,
              use_raster = TRUE, raster_by_magick = TRUE)



ht_list = hm1+hm2+anno_txt

draw(ht_list,
     row_title = "dAUC",
     row_title_gp = gpar(col = "black"),
     heatmap_legend_side = "right",
     annotation_legend_side = "right",
     legend_labels_gp = gpar(24))
```
```{r}
ht_list = rAnno+anno_txt
draw(ht_list)
```


```{r}
tmp = cbind(anno_df,Rx_top,Mid_v_Pre,Pst_v_Pre)
write.csv(tmp,"20220920 Differential drug actvities.csv")
```


```{r}
hm1 = Heatmap(t(Mid_v_Pre),
              name = "dAUC",
              clustering_method_column = "ward.D2"
              #column_title = "postNACT-preNACT",
              #column_title_gp=gpar(fontsize=20,fontface = "bold"),
              #row_split = anno_df$Class,
              #row_title_rot = 0,
              #row_title_gp=gpar(fontsize=24,fontface = "bold"),
              #heatmap_legend_param = list(direction = "vertical")
              )

hm2 = Heatmap(t(Pst_v_Pre),
              name = "dAUC"
              #column_title = "Surgery-preNACT",
              #column_title_gp=gpar(fontsize=20,fontface = "bold"),
              #heatmap_legend_param = list(direction = "vertical")
              )

ht_list = hm1+hm2#+rAnno

draw(ht_list,
     row_title = "dAUC",
     row_title_gp = gpar(col = "black"),
     #column_title = "Differential drug susceptability",
     #column_title_gp = gpar(fontsize = 24),
     heatmap_legend_side = "right",
     annotation_legend_side = "right")
```


```{r}
Y_all = t(Drug_all[drug_filter,train_samples])
drugnames = colnames(Y_all)

#X_all = t(RNA_filtered[,common])
X_all = t(RNA_filtered_RWR[,train_samples])
```


```{r}
library(mRMRe)
featureSelection <- function(x, y, method, features.no, shrink=T, x_raw=NULL){
  if(is.null(x_raw)){
    x_raw <- x
  }
  switch(method,
         "mRMR"={
           if(shrink){
             var_features <- apply(x_raw, MARGIN=2, sd, na.rm=T)
             mad_features <- apply(x_raw, MARGIN=2, mad, na.rm=T)
             temp <- which(var_features > quantile(var_features, .75, na.rm=T) & 
                             mad_features > quantile(mad_features, .75, na.rm=T))
             if(length(temp) > features.no){
               x <- x[, temp, drop=FALSE]
             }
           }
           
           f_data <- mRMR.data(data=as.data.frame(cbind(x, y), stringAsFactor=FALSE))
           features <- mRMR.ensemble(data=f_data,
                                     target_indices=ncol(x) + 1,
                                     feature_count=features.no,
                                     solution_count=1)
           features <- features@feature_names[unlist(features@filters)]
         },
         "variance"={
           var_features <- apply(x_raw, MARGIN=2, var, na.rm=T)
           features <- names(sort(var_features, decreasing=T))[1:features.no]
         },
         "mad"={
           var_features <- apply(x_raw, MARGIN=2, mad, na.rm=T)
           features <- names(sort(var_features, decreasing=T))[1:features.no]
         }
         ,
         "r_sq"={
           var_features <- cor(x_raw,y,method="pearson")^2
           features <- row.names(var_features[order(var_features[,1], decreasing = T),,drop=F])[1:features.no]
         })
  return(features)
}
```

Random forest trained with Leave-one-out-cross-validation
```{r}
library("caret")
library("ranger")
library("randomForest")
library("doParallel")
cl <- makePSOCKcluster(30)
registerDoParallel(cl)

Model_out = "MODEL_20220916"

Model_out = file.path(Base_dir,Model_out)
save_importance = file.path(Model_out, "Importance")
save_QC = file.path(Model_out, "QC")
save_binary = file.path(Model_out, "bin")

dir.create(Model_out, showWarnings = FALSE)
dir.create(save_importance, showWarnings = FALSE)
dir.create(save_binary, showWarnings = FALSE)
dir.create(save_QC, showWarnings = FALSE)

Act_v_Pred_boot = c()
Act_v_Pred_withheld = c()

for (active_drug in 1:ncol(Y_all)){
feat = featureSelection(x=X_all,
             y=Y_all[,active_drug],
             method="r_sq",
             features.no = 2000,
             shrink = T)
  
  df = as.data.frame(cbind(Cmpd=Y_all[,active_drug],X_all[,feat]))
  active_drug_name = drugnames[active_drug]
  
  set.seed(1536)
  # Fit random forest: model
  model <- train(
    Cmpd ~.,
    tuneLength = 10,
    data = df, 
    method = "ranger",
    importance = 'permutation',
    metric = "RMSE",
    trControl = trainControl(method = "LOOCV",
                             number = 1,
                             verboseIter = FALSE))
    
    #Get iteratative performance data
    predicted <- unlist(model[["finalModel"]][["predictions"]])
    actual = df$Cmpd

    #Actual vs predicted from model bootstrap
    temp = cbind(active_drug_name,actual,predicted)
    row.names(temp) = train_samples
    Act_v_Pred_boot = rbind(Act_v_Pred_boot,temp)

    #Actual vs predicted from the with-held dataset
    act = unlist(as.list(with_held_Drug[active_drug_name,]))
    final_gene_names = colnames(X_all)
    input = t(with_held_RNA[final_gene_names,])
    pred = predict(model,input)
    temp = cbind(active_drug_name,act,pred,interval="confidence", level=0.90, type="response")
    Act_v_Pred_withheld = rbind(Act_v_Pred_withheld,temp)
    
    #Export RANGER importance
    imp_p = importance_pvalues(model$finalModel,method=c("janitza","altmann"),num.permutations = 100)
    temp = cbind(varImp(model)[["importance"]],imp_p)
    write.csv(temp,
              file=paste(save_importance,"/",active_drug_name,".csv",sep=""),
              row.names=TRUE)
    
    #Save model
    save(model, file=paste(save_binary,"/",active_drug_name,".RData",sep=""))
}

stopCluster(cl)

flnm = 'Act_v_Pred_cv.csv'
write.csv(Act_v_Pred_boot,
          file=paste(save_QC,flnm,sep="/"))

flnm = 'Act_v_Pred_withheld.csv'
write.csv(Act_v_Pred_withheld,
          file=paste(save_QC,flnm,sep="/"))
```

Heatmaps/materials for PGx figure
```{r}
Heatmap(RNA_all[,train_samples],
        show_row_names = F,
        show_row_dend = F,
        #cluster_rows = F,
        #cluster_columns = F,
        show_column_names = F,
        show_column_dend = F)
```

```{r}
Heatmap(RNA_filtered_RWR[,train_samples],
        show_row_names = F,
        show_row_dend = F,
        #cluster_rows = F,
        #cluster_columns = F,
        show_column_names = F,
        show_column_dend = F)
```

```{r}
Heatmap(Drug_all[drug_filter,train_samples],
        show_row_names = F,
        show_row_dend = F,
        #cluster_rows = F,
        #cluster_columns = F,
        show_column_names = F,
        show_column_dend = F)
```


Actual_v_predicted for bootstrapped sample
```{r}
temp = transform(Act_v_Pred_boot,
                 actual = as.numeric(as.character(actual)),
                 predicted = as.numeric(as.character(predicted))
                 )
cor.test(temp$actual,temp$predicted,method="pearson")

temp %>%
  ggplot(aes(x=actual,y=predicted))+
  geom_point()+
  geom_smooth(method="lm")+
  theme(text = element_text(size = 24))
```

Actuval_v_predicted for all samples
```{r}
temp = transform(Act_v_Pred_withheld,
                 actual = as.numeric(as.character(act)),
                 predicted = as.numeric(as.character(pred))
                 )

cor(temp$actual,temp$predicted,method="pearson")

temp %>%
  ggplot(aes(x=actual,y=predicted))+
  geom_point()+
  geom_smooth(method="lm")+
  theme(text = element_text(size = 24))
```

```{r}
interactive_network = function(result_df,save_as){
  library("visNetwork")
  
  result_df <- result_df[order(result_df$lowest_p, decreasing = FALSE), ]
  
  ID_column <- "Term_Description"
  
  edges <- data.frame()
  all_up = c()
  all_down = c()
  for (i in base::seq_len(nrow(result_df))) {
    up_genes <- unlist(strsplit(result_df$Up_regulated[i],", "))
    all_up = unique(c(all_up,up_genes))
    down_genes <- unlist(strsplit(result_df$Down_regulated[i],", "))
    all_down = unique(c(all_down,down_genes))
    genes <- c(up_genes, down_genes)
    for (gene in genes) {
      edges <- rbind(edges, data.frame(Term = result_df[i, ID_column], Gene = gene))
    }
  }
  
  pathways = as.data.frame(unique(edges$Term))
  colnames(pathways) = "id"
  pathways[,"color"] = "grey"
  pathways[,"group"] = "Pathway"
  pathways[,"shape"] = "square"
  
  all_up = as.data.frame(all_up)
  colnames(all_up) = "id"
  all_up[,"color"] = "green"
  all_up[,"group"] = "Up_regulated"
  all_up[,"shape"] = "circle"
  
  # all_down = as.data.frame(all_down)
  # colnames(all_down) = "id"
  # all_down[,"color"] = "darkred"
  # all_down[,"group"] = "Down_regulated"
  # all_down[,"shape"] = "circle"
  
  #nodes = rbind(pathways,all_up,all_down)
  nodes = rbind(pathways,all_up)
  
  colnames(edges) = c("from","to")
  
  rm("pathways","all_up","all_down")
  
  graph = visNetwork(nodes, edges,width='100%',height=2000)%>%
    visNodes(color = list(inherit=FALSE))%>%
    visEdges(color = list(inherit=FALSE))%>%
    visOptions(highlightNearest = list(enabled = T, degree = 2, hover = T)) %>%
    visInteraction(multiselect = TRUE) %>%
    visLayout(randomSeed = 123)%>%
    visGroups(groupname = "Pathway", color = "grey") %>%
    visGroups(groupname = "Up_regulated", color = "green") %>%
    visGroups(groupname = "Down_regulated", color = "darkred") %>%
    visLegend(position = "right")
  visSave(graph, file=save_as, selfcontained = TRUE, background = "white")
}
```



```{r message=FALSE, warning=FALSE, include=FALSE}
library("stringr")
library("pathfindR")

pathfindR_drug_list = list()

files = list.files(path="./MODEL_20220916/Importance/",
           pattern = ".csv",
           full.names = TRUE)

for(file in files){
  flnm = str_replace(str_remove(basename(file),".csv"),' ','')
  save_as = paste("./MODEL_20220916/pathfindR",flnm,sep="/")

  drug_of_interest <- read.csv(file)
  drug_of_interest = drug_of_interest[,c("X","Overall","pvalue")]
  
  unlink(save_as, recursive = TRUE)

  tmp = run_pathfindR(drug_of_interest,iterations = 10,
                      p_val_threshold = 0.05,
                      search_method = "GR", #"GR" = Greedy algorythm (Default). "SA" = Simulated annealing, "GA" = Genetic algorythm
                      plot_enrichment_chart = FALSE,
                      visualize_enriched_terms = FALSE,
                      output_dir = save_as,
                      silent_option = TRUE)
  interactive_network(tmp,paste("./MODEL_20220916/pathfindR/",flnm,"/visNetork.html",sep=""))
  pathfindR_drug_list[[flnm]] = tmp
}
```

```{r}
selected_drug = pathfindR_drug_list[["PEVONEDISTAT_Broad"]]
selected_drug
```

```{r fig.height=5, fig.width=12}
enrichment_chart(
  selected_drug,
  top_terms = 15,
  plot_by_cluster = TRUE,
  num_bubbles = 5,
  even_breaks = TRUE
)
```
```{r}
DEG_pathways = DEG_pathfindR$Term_Description
PGx_pathways = selected_drug$Term_Description

#get list of correlated and variable drug names
x = list(DEG_pathways = DEG_pathways,
         PGx_pathways = PGx_pathways)

library(ggvenn)
ggvenn(x,
       fill_color = c("#0073C2FF","#CD534CFF"),
       stroke_size = 0.5, set_name_size = 4)
```


```{r}
DEG_pathways = row.names(DEG_top[which(DEG_top$P.Value <= 0.05),])
  
result_df = selected_drug
result_df <- result_df[order(result_df$lowest_p, decreasing = FALSE), ]

ID_column <- "Term_Description"

edges <- data.frame()
all_up = c()
all_down = c()
for (i in base::seq_len(nrow(result_df))) {
  up_genes <- unlist(strsplit(result_df$Up_regulated[i],", "))
  all_up = unique(c(all_up,up_genes))
  down_genes <- unlist(strsplit(result_df$Down_regulated[i],", "))
  all_down = unique(c(all_down,down_genes))
  genes <- c(up_genes, down_genes)
  for (gene in genes) {
    edges <- rbind(edges, data.frame(Term = result_df[i, ID_column], Gene = gene))
  }
}

PGx_pathways = all_up

#get list of correlated and variable drug names
x = list(DEG_pathways = DEG_pathways,
         PGx_pathways = PGx_pathways)

library(ggvenn)
ggvenn(x,
       fill_color = c("#0073C2FF","#CD534CFF"),
       stroke_size = 0.5, set_name_size = 4)
```



```{r}
intersect(DEG_pathways,PGx_pathways)
```

```{r fig.height=6, fig.width=14}
term_gene_heatmap(selected_drug,
                  use_description = TRUE,
                  num_terms = 25,
                  low = "red",
                  mid = "black",
                  high = "green")
```


```{r fig.height=8, fig.width=20}
term_gene_graph(selected_drug,
                use_description = TRUE,
                num_terms = 25,
                node_size = "p_val",
                layout = "stress")
```


```{r message=FALSE, warning=FALSE}
library("dplyr")
library("edgeR")

sel = c("PIM137","PIM172",
        "PIM025","PIM038",
        "PIM190","PIM231",
        "PIM254","PIM311")
PEV_ = c("RESP","NON",
          "NON","NON",
          "NON","NON",
          "RESP","RESP")
RNA_select = RNA_all %>% dplyr::select(sel)
Meta_select = Meta_all[sel,]
Meta_select = cbind(Meta_select,PEV_)
Meta_select[] <- lapply(Meta_select, factor)

design = model.matrix(~0+PEV_+PiD,Meta_select)
PEV_DEG_fit = lmFit(RNA_select,design)
cm = makeContrasts(PEV_RESP=PEV_RESP-PEV_NON,levels=design)
PEV_DEG_fit_cm = contrasts.fit(PEV_DEG_fit,cm)
PEV_DEG_fit_cm_eBayesfit = eBayes(PEV_DEG_fit_cm,trend = FALSE,robust = FALSE)
PEV_DEG_top = topTable(PEV_DEG_fit_cm_eBayesfit,adjust.method="BH", number=Inf)

tmp = PEV_DEG_top[which(PEV_DEG_top$P.Value <= 0.05),]
up = as.integer(length(which(tmp$logFC > 0)))
down =  as.integer(length(which(tmp$logFC <= 0)))
print(paste("Total DEG: ", up+down))
print(paste("Sigificantly up-reglated genes: ", up))
print(paste("Sigificantly down-reglated genes: ", down))

PEV_DEG_top

plot(density(PEV_DEG_top$P.Value))

library("plotly")
fig <- plot_ly(data = PEV_DEG_top, x = ~logFC, y = ~-1*log10(P.Value),color = ~logFC, colors="RdBu")%>%
  layout(xaxis = list(title = "log(FC)"),yaxis = list(title = "-log(p-value)")) %>%
  add_trace(text=row.names(PEV_DEG_top))
fig

rm("tmp","up","down")
```


```{r}
# install.packages("devtools") # if you have not installed "devtools"
# devtools::install_github("egeulgen/pathfindR")

library("pathfindR")
pfR_input = PEV_DEG_top[,c("logFC","P.Value"),]
Gene_symbol = row.names(pfR_input)
pfR_input = cbind(Gene_symbol,pfR_input)

unlink("InVivo_PEV_DEG_pathfindR", recursive = TRUE)

PEV_DEG_pathfindR <- run_pathfindR(pfR_input,
                               iterations = 10,
                               p_val_threshold = 0.05,
                               search_method = "GR",
                               plot_enrichment_chart = FALSE,
                               visualize_enriched_terms = FALSE,
                               output_dir = "InVivo_PEV_DEG_pathfindR",
                               silent_option = TRUE)
PEV_DEG_pathfindR
```

```{r}
library("visNetwork")

result_df = PEV_DEG_pathfindR
result_df <- result_df[order(result_df$lowest_p, decreasing = FALSE), ]

ID_column <- "Term_Description"

edges <- data.frame()
all_up = c()
all_down = c()
for (i in base::seq_len(nrow(result_df))) {
  up_genes <- unlist(strsplit(result_df$Up_regulated[i],", "))
  all_up = unique(c(all_up,up_genes))
  down_genes <- unlist(strsplit(result_df$Down_regulated[i],", "))
  all_down = unique(c(all_down,down_genes))
  genes <- c(up_genes, down_genes)
  for (gene in genes) {
    edges <- rbind(edges, data.frame(Term = result_df[i, ID_column], Gene = gene))
  }
}

pathways = as.data.frame(unique(edges$Term))
colnames(pathways) = "id"
pathways[,"color"] = "grey"
pathways[,"group"] = "Pathway"
pathways[,"shape"] = "square"

all_up = as.data.frame(all_up)
colnames(all_up) = "id"
all_up[,"color"] = "green"
all_up[,"group"] = "Up_regulated"
all_up[,"shape"] = "circle"

all_down = as.data.frame(all_down)
colnames(all_down) = "id"
all_down[,"color"] = "darkred"
all_down[,"group"] = "Down_regulated"
all_down[,"shape"] = "circle"

nodes = rbind(pathways,all_up,all_down)

colnames(edges) = c("from","to")

rm("pathways","all_up","all_down")

graph = visNetwork(nodes, edges,width='100%',height=2000)%>%
  visNodes(color = list(inherit=FALSE))%>%
  visEdges(color = list(inherit=FALSE))%>%
  visOptions(highlightNearest = list(enabled = T, degree = 2, hover = T)) %>%
  visInteraction(multiselect = TRUE) %>%
  visLayout(randomSeed = 123)%>%
  visGroups(groupname = "Pathway", color = "grey") %>%
  visGroups(groupname = "Up_regulated", color = "green") %>%
  visGroups(groupname = "Down_regulated", color = "darkred") %>%
  visLegend(position = "right")
visSave(graph, file="InVivo_PEV_DEG_pathfindR.html", selfcontained = TRUE, background = "white")
graph
```


```{r}
save.image(file='TNBC-PGx_analysis.RData')
sessionInfo()
```
